Enabling GPU Virtualization in Cloud Environments

### 2015-10-26 paper.tex
intro re-revisado

\section{\uppercase{Introduction}}
\label{sec:introduction}
Nowadays, many cloud vendors have started offering virtual machines (VMs) with GPUs in order to provide
GPGPU computation services to users. A few relevant examples include 
Amazon Web Services (AWS)\footnote{\url{https://aws.amazon.com}}, 
Penguin Computing\footnote{\url{http://www.penguincomputing.com}}, 
Softlayer\footnote{\url{http://www.softlayer.com}} and 
Microsoft Azure\footnote{\url{https://azure.microsoft.com}}.
In the public scope, one of the most popular cloud vendors is AWS, that offers a wide range of preconfigured instances ready to be launched.
Alternatively, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
Openstack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.

Unfortunately, sharing GPU resources among multiple VMs in cloud environments 
is not straightforward as is in a physical servers resulting into a low utilization rate.
On one hand, instances in public clouds are not easily customizable. 
On the other, in a private cloud, although the instances can be customized in many aspects, when referring to GPUs
the number of options is reduced.
Moreover, neither vendors nor tools are offer GPGPU wasting the opportunity of providing flexibility and boosty the utilization of GPU resources in both private or public clouds.

These devices were adopted with the release of the CUDA programming model for
NVIDIA GPUs~\cite{cuda65}, which includes both high- and low-level application programming interfaces (APIs)
to allow the use of NVIDIA devices as general-purpose
accelerators.
Following CUDA, OpenCL~\cite{opencl} appeared as an attempt
to offer a cross-vendor solution.

The current approach to assemble a cluster with accelerators consist in installing 
one or more of these devices in each compute node.
However, these configurations often present a low
utilization rate of the computational resources
 because of mismatches between the application's type of parallelism
and the accelerators' architecture.

Remote virtualization has been recently proposed to deal with the low-usage problem 
(e.g. {rCUDA}~\cite{tonithesis},
DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}).

Roughtly speaking these virtualization frameworks 
enable cluster configurations with fewer GPUs than nodes.  The goal is that 
GPU-equipped nodes act as GPGPU (general-purpose GPU) servers, yielding a CUDA-sharing solution that potentially achieves
a higher overall GPU load in the system.

The main goals of this work are to study current cloud solutions from 
the point of view of HPC field and GPU usage, and to analyze and improve 
them by adding flexibility with GPU virtualization. In order to reach 
this goal, we select rCUDA, a virtualization tool that is possibly the more complete 
and up-to-date for NVIDIA GPUs.

The rest of the paper is structured as follows. 
In Section~\ref{sec:background} the technologies we introduce used in this work;
Section~\ref{sec:state} summarizes previous work in this field; 
Section~\ref{sec:motivation} states the motivation to carry out this work;
the effort to use AWS is explained in Section~\ref{sec:workingaws};
while the work with Openstack is described in Section~\ref{sec:gpgpuOS};
finally Section~\ref{sec:conclusions} summarizes the advances and Section~\ref{sec:future} proposes the next steps
 of this research.

>>>>

\section{\uppercase{Introduction}}
\label{sec:introduction}
Nowadays, many cloud vendors have started offering virtual machines (VMs) with graphics processing units (GPUs) in order to provide
GPGPU (general-purpose GPU) computation services. A few relevant examples include 
Amazon Web Services (AWS)\footnote{\url{https://aws.amazon.com}}, 
Penguin Computing\footnote{\url{http://www.penguincomputing.com}}, 
Softlayer\footnote{\url{http://www.softlayer.com}} and 
Microsoft Azure\footnote{\url{https://azure.microsoft.com}}.
In the public scope, one of the most popular cloud vendors is AWS, which offers a wide range of preconfigured instances ready to be launched.
Alternatively, owning the propper infrastructure, a private cloud can be deployed using a specific middleware such as
Openstack\footnote{\url{https://www.openstack.org}} or Opennebula\footnote{\url{http://opennebula.org}}.

Unfortunately, sharing GPU resources among multiple VMs in cloud environments 
is as quite more complex than in physical servers, resulting in a low utilization rate.
On one hand, instances in public clouds are not easily customizable. 
On the other, in a private cloud, although the instances can be customized in many aspects, when referring to GPUs
the number of options is reduced.
Moreover, neither vendors nor tools offer GPGPU services, missing the opportunity of providing flexibility and boost the utilization of GPU resources in both private or public clouds.

Remote virtualization has been recently proposed to deal with the low-usage problem. Some relevant examples include {rCUDA}~\cite{tonithesis},
DS-CUDA~\cite{dscuda}, gVirtus~\cite{gvirtus}, vCUDA~\cite{vcuda}, VOCL~\cite{vocl}, and SnuCL~\cite{snucl}.

Roughtly speaking these virtualization frameworks 
enable cluster configurations with fewer GPUs than nodes.  The goal is that 
GPU-equipped nodes act as GPGPU servers, yielding a GPU-sharing solution that potentially achieves
a higher overall utilization of the accelerators in the system.

The main goals of this work are to study current cloud solutions in an HPC GPU-enabled scenario, and to analyze and improve 
them by adding flexibility with GPU virtualization. In order to reach 
this goal, we select rCUDA, a virtualization tool that is possibly the more complete 
and up-to-date for NVIDIA GPUs.

The rest of the paper is structured as follows. 
In Section~\ref{sec:background} we introduce the technologies used in this work;
Section~\ref{sec:state} summarizes related work; 
Section~\ref{sec:motivation} states the motivation to address this task;
the effort to use AWS is explained in Section~\ref{sec:workingaws};
while the work with Openstack is described in Section~\ref{sec:gpgpuOS};
finally, Section~\ref{sec:conclusions} summarizes the advances and Section~\ref{sec:future} outlines the next steps
 of this research.
